---
title: "class09"
Author: Khine Win 
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Unsupervised Learning Analysis of Cancer Cells

import from the data, Values in this data set describe characteristics of the cell nuclei present in digitized images of a fine needle aspiration (FNA) of a breast mass. For example radius (i.e. mean of distances from center to points on the perimeter), texture (i.e. standard deviation of gray-scale values), and smoothness (local variation in radius lengths). Summary information is also provided for each group of cells including diagnosis (i.e. benign (not cancerous) and and malignant (cancerous)).
```{r}
url <- "https://bioboot.github.io/bimm143_W18/class-material/WisconsinCancer.csv"
wisc.df  <- read.csv(url)
head(wisc.df)

```
so, the diagonosis col is the kabels , let's drop this and the id col so we just have the measurement as a matrix, take a colum
## Convert the features of the data: wisc.data

```{r}

wisc.data  <- as.matrix(wisc.df[,3:32], ncol=30)
wisc.data
```

# Set the row names of wisc.data
```{r}
row.names(wisc.data)  <- wisc.df$id
```
```{r}
table(wisc.df$diagnosis)
```

setup a separate new vector called diagnosis to be 1 if a diagnosis is malignant ("M") and 0 otherwise. Note that R coerces TRUE to 1 and FALSE to 0.


# Create diagnosis vector by completing the missing code
```{r}
diagnosis   <-  as.numeric(wisc.df$diagnosis  == "M")
sum(diagnosis)
```

#Q1. How many observations are in this dataset?
```{r}
dim(wisc.data)

```
Q2. How many variables/features in the data are suffixed with _mean?
```{r}

x  <- length(grep("_mean",colnames(wisc.data)))
```
There are `r x` means measurement im the data 

Q3. How many of the observations have a malignant diagnosis?

```{r}
sum(x, na.rm = FALSE)
```



# Section 2 ,Performing PCA
## Check column means and standard deviations

``colMeans(wisc.d`{r}
```{r}
colMeans(wisc.data)
apply(wisc.data,2,sd)
```


# Perform PCA on wisc.data by completing the following code

```{r}
wisc.pr <- prcomp(wisc.data, scale= TRUE)

```

# Look at summary of results
```{r}
summary(wisc.pr)
```

let's make plot PC1 VS PC2
```{r}
attributes(wisc.pr)
```

```{r}
dim(wisc.pr$x)
```

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2],col=diagnosis+1)
```

##Interpreting PCA results

```{r}
biplot(wisc.pr)
```


#Variance explained
# Variance explained by each principal component: pve

```{r}
pr.var <- wisc.pr$sdev^2
pve  <-   pr.var/ sum(pr.var)
```

```{r}
# Plot variance explained for each principal component
plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```


# Alternative scree plot of the same data, note data driven y-axis

```{r}
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```


# Plot cumulative proportion of variance explained( using cumsum)
```{r}
plot(cumsum(pve), xlab = "Principal Component", 
     ylab = "Cumulative Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```
#Q9. For the first principal component, what is the component of the loading vector (i.e.  wisc.pr$rotation[,1]) for the feature concave.points_mean?


#Q10. What is the minimum number of principal components required to explain 80% of the variance of the data?



##Section 3.
Hierarchical clustering of case data

# Scale the wisc.data data: data.scaled
```{r}
data.scaled <- scale(wisc.data)

```


we need a distance matix for hierarchical clustering input 
```{r}
data.dist  <- dist(data.scaled)
```


# Create a hierarchical clustering model using complete linkage. Manually specify the method argument to hclust() and assign the results to wisc.hclust.

```{r}
wisc.hclust <- hclust(data.dist,method = "complete")
```
# Results of hierarchical clustering
```{r}
plot(wisc.hclust)
```

#Selecting number of clusters

```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, k=4)
table(wisc.hclust.clusters)
```
#compare cluster groupp to our digonosis 

```{r}
table(wisc.hclust.clusters, diagnosis)
```

#Q12. Can you find a better cluster vs diagnoses match with by cutting into a different number of clusters between 2 and 10?

##Section 4

##Section 5.
Clustering on PCA results

### Use the distance along the first 7 PCs for clustering i.e. wisc.pr$x[, 1:7]
```{r}
d.pr <- dist(wisc.pr$x[, 1:7])
wisc.pr.hclust <- hclust(d.pr, method="complete")
plot(wisc.pr.hclust)
```


```{r}
wisc.pr.hclust.clusters <- cutree(wisc.pr.hclust, k=4)
plot(wisc.hclust.clusters)

```

# Compare to actual diagnoses
#Q14. How well does the newly created model with four clusters separate out the two diagnoses?

```{r}
table(wisc.hclust.clusters,diagnosis)
```


#Q15. How well do the k-means and hierarchical clustering models you created in previous sections (i.e. before PCA) do in terms of separating the diagnoses? Again, use the table() function to compare the output of each model (wisc.km$cluster and wisc.hclust.clusters) with the vector containing the actual diagnoses.


## BONUS Predictive PC Modeling
take a new patient data and aooly to out PCA model from 
## Predicting Malignancy Of New samples 

```{r}
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
plot(wisc.pr$x[,1:2], col=diagnosis+1)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
```












