---
title: "Class 8"
author: Khine Win
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## K-means clustering

Our first example with **kmean()** function.

```{r}
# Example plot to see how Rmarkdown works
plot(1:10,typ="l")
```


back to kmean ...

```{r}
# Generate some example data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
```

Use the kmeans() function setting k to 2 and nstart=20


 Inspect/print the results
 
Q. How many points are in each cluster?


Q. What ‘component’ of your result object details- 
-cluster size?

 - cluster assignment/membership?
 - cluster center?
 Plot x colored by the kmeans cluster assignment and
 add cluster centers as blue points
 Q. Repeat for k=3, which has the lower tot.withinss? 

```{r}
k  <- kmeans(x, centers=2, nstart= 20)
k
```

Inspect/print the results
 
Q. How many points are in each cluster?


Q. What ‘component’ of your result object details- 
-cluster size?
```{r}
k$size
```


 - cluster assignment/membership?
```{r}
k$cluster
```

```{r}
table(k$cluster)
```
 
 - cluster center?
```{r}
k$centers
```
 
 Plot x colored by the kmeans cluster assignment and
 add cluster centers as blue points
```{r}

palette(c("blue","red"))
plot(x, col=k$cluster)
points(k$centers, col="green",pch=20, cex=3)
```
 
 Q. Repeat for k=3, which has the lower tot.withinss? 
```{r}
k2 <- kmeans(x, centers = 2, nstart = 20)
k3 <- kmeans(x, centers = 3, nstart = 20)
k4 <- kmeans(x, centers = 4, nstart = 20)
k5 <- kmeans(x, centers = 5, nstart = 20)

k2$tot.withinss
k3$tot.withinss
k4$tot.withinss
k5$tot.withinss

plot(c(k2$tot.withinss,k3$tot.withinss,k4$tot.withinss,k5$tot.withinss), type = "b")
```
 
```{r}
k3  <- kmeans(x, centers = 3, nstart= 20)
k3$tot.withinss
k$tot.withinss

```
 
##Hierarchical clustering in R 

let's try out the **hclust()**fcuntions for Hierarchical clustering in R 
```{r}
d <- dist(x)
hc <- hclust(d)
plot(hc)
```

```{r}
plot(hc)
# draw a line on the dendrogram
abline(h=8, col="red")

# cut the tree to yeild cluster membership vector
cutree(hc, h=8) 
cutree(hc, k=2)
```


### A more "real" example of data clustering 
```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
# Step 2. Plot the data without clustering
plot(x)
# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )

grp.names  <- rep(c("red","blue","black"), each=50)
col <- as.factor(rep(c("red","blue","black"), each=50))

palette(c("red","blue","darkgreen"))
plot(x, col=col,pch=20)
```

 Q. Use the dist(), hclust(), plot() and cutree()
functions to return 2 and 3 clusters
```{r}
d  <- dist(x)
hc  <- hclust(d)
plot(hc)
```
cutree()
functions to return 2 and 3 clusters

```{r}
grp2  <- cutree(hc,k=2)
grp3   <- cutree(hc,k=3)
plot(x, col=grp3, pch=20)
```

 Q. How does this compare to your known 'col' groups?
```{r}
# use the table function !
table(grp3,col)
```
 


# princial component analysis 

lets make up some data with quite a few dimension 

```{r}
## Initialize a blank 100 row by 10 column matrix
mydata <- matrix(nrow=100, ncol=10)
## Lets label the rows gene1, gene2 etc. to gene100
rownames(mydata) <- paste("gene", 1:100, sep="")
## Lets label the first 5 columns wt1, wt2, wt3, wt4 and wt5
## and the last 5 ko1, ko2 etc. to ko5 (for "knock-out")
colnames(mydata) <- c( paste("wt", 1:5, sep=""),
 paste("ko", 1:5, sep="") )
## Fill in some fake read counts
for(i in 1:nrow(mydata)) {
 wt.values <- rpois(5, lambda=sample(x=10:1000, size=1))
 ko.values <- rpois(5, lambda=sample(x=10:1000, size=1))

 mydata[i,] <- c(wt.values, ko.values)
}
head(mydata)
```




```{r}
head(t(mydata))
```


Now lets try to find "structure"" in this data with **prcomp()**
```{r}
pca <- prcomp(t(mydata), scale=TRUE)
pca
```



make a PC plot of PC1 vs PC2 for this I use the '$x' component of our output. (i.e, pca$x).

```{r}
## A basic PC1 vs PC2 2-D plot
plot(pca$x[,1], pca$x[,2]) 
```




let;s see how well our PC are doing (i,e how much variance form the original data are they capatureing )

```{r}
## Precent variance is often more informative to look at
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)

pca.var.per
```



Make a "screen plot"" to see the variance captured in each PC
```{r}
barplot(pca.var.per, main="Scree Plot",
 xlab="Principal Component", ylab="Percent Variation")
```



lets make our PC more useful...
```{r}
## A vector of colors for wt and ko samples
colvec <- colnames(mydata)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
 xlab=paste0("PC1 (", pca.var.per[1], "%)"),
 ylab=paste0("PC2 (", pca.var.per[2], "%)")) 

```





```{r}
x <- read.csv("UK_foods.csv")
```












