---
title: "class 13: Structure Bioinformatics II"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Get HIV pr structure from PDB databases 
we will work with the structure '1HSG'
```{r}
library(bio3d)
file.name <- get.pdb("1hsg")
```

Read this into R and examine composition 
```{r}
hiv  <- read.pdb(file.name)
hiv
```


## slit into separate protein and ligand files 

We will use the 'trim.pdb()' function to split our input structure.
```{r}
prot <- trim.pdb(hiv, "protein")
prot
```


```{r}
write.pdb(prot,file = "1hsg_protein.pdb")
```

```{r}
lig <- trim.pdb(hiv, "ligand")
lig
```

```{r}
write.pdb(lig, "1hsg_ligand.pdb")
```

## blocking with Vina
We run this commads:
'~/Downloads/autodock_vina_1_1_2_mac/bin/vina --config config.txt --log log.txt'


## Readd docking reuslts
we got a file all.pdbqt that we need to make nito a PDB format

```{r}
res <- read.pdb("all.pdbqt", multi=TRUE)
write.pdb(res, "results.pdb")
```

```{r}
res
```

To assess the results quantitatively we will calculate the RMSD (root mean square distance)
between each of the docking results and the known crystal structure using the bio3d package.
Back in RStudio read the original ligand with added hydrogens that you produced earlier and
use the rmsd() function to compare to your docking results.

```{r}
ori <- read.pdb("1hsg_ligand.pdbqt")
rmsd(ori, res)
```


























